---
title: "Simple registration examples"
author: "Lars Lau Raket"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simple examples}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, comment = "#>", fig.height = 5, fig.width = 7)
library(pavpop)
```
In this vignette we will consider some simple examples of data that needs registration.

## Example 1: linear shift

First, we generate a dataset with randomly shifted curves with noise
```{r}
# Number of samples
n <- 30 
# Number of observation points
m <- 30 

# Observation points
t <- seq(0, 1, length = m + 2)[2:(m + 1)] 

# Mean function
theta <- function(t) dnorm(t, mean = 0.3, sd = 0.2) + dnorm(t, mean = 0.7, sd = 0.15)

# Generate shifts
w <- runif(n, min = -0.1, max = 0.1)

# Generate data with random shifts
sigma <- 0.1
y <- lapply(w, function(w) {theta(t + w) + rnorm(m, sd = sigma)}) 
t <- lapply(1:n, function(x) t)

# Plot shifted curves
plot(0, 0, xlim = c(-0.2, 1.2), ylim = range(y), type = 'n', 
     xlab = 't', ylab = 'y(t)')
legend(0.9, 3, legend = expression(theta(t)), lty = 2, lwd = 2)
for (i in 1:n) lines(t[[i]], y[[i]], col = rainbow(n)[i])
lines(t[[1]], theta(t[[1]]), lwd = 2, lty = 2)
```

We now set up the pavpop model to estimate in the model

```{r}
# Set up basis function
kts <- seq(-0.2, 1.2, length = 15)
basis_fct <- make_basis_fct(kts = kts, intercept = TRUE, order = 3, boundary = c(-0.2, 1.2))

# Set up warp function
warp_fct <- make_warp_fct(type = 'shift')

# Estimate in the model
res <- pavpop(y, t, basis_fct, warp_fct, amp_cov = NULL, warp_cov = NULL, iter = c(1, 5))
```

We can not plot the results

```{r}
plot(w, res$w, xlab = 'true shifts', ylab = 'estimated shifts', pch = 19)
abline(0, 1, lty = 2)

t_plot <- seq(-0.2, 1.2, length = 100)
plot(0, 0, xlim = c(-0.2, 1.2), ylim = range(y), type = 'n', xlab = 't', ylab = expression(theta(t)))
for (i in 1:n) lines(t[[i]] + res$w[i], y[[i]], col = rainbow(n)[i])
lines(t_plot, basis_fct(t_plot) %*% res$c, lwd = 2, lty = 2, col = 'red')
lines(t_plot, theta(t_plot), lwd = 2, lty = 2)
```


## Example 2: linear shift with normally distributed shifts
We use the same data as before, but now we generate random shifts from a normal distribution
```{r}
# Generate shifts
scale <- 0.5
w <- rnorm(n, sd = sigma * sqrt(scale))

# Generate data with random shifts
y <- lapply(1:n, function(i) {theta(t[[i]] + w[i]) + rnorm(m, sd = sigma)}) 

# Plot shifted curves
plot(0, 0, xlim = c(-0.2, 1.2), ylim = range(y), type = 'n', xlab = 't', ylab = expression(theta(t)))
for (i in 1:n) lines(t[[i]], y[[i]], col = rainbow(n)[i])
lines(t[[1]], theta(t[[1]]), lwd = 2, lty = 2)
```

Let us first ignore the uncertainty and use the same model as in Example 1.

```{r}
res_no_uncert <- pavpop(y, t, basis_fct, warp_fct, amp_cov = NULL, warp_cov = NULL, iter = c(1, 5))
```

Correct specification of the model where the shifts are a normal random effect can be done as follows

```{r}
warp_cov <- make_cov_fct(id_cov)
res <- pavpop(y, t, basis_fct, warp_fct, amp_cov = NULL, warp_cov = warp_cov, iter = c(5, 5))
```


We can not plot the results

```{r}
plot(w, res$w, xlab = 'true shifts', ylab = 'estimated shifts', pch = 19, col = 'green')
points(w, res_no_uncert$w, pch = 19, col = 'red', cex = 0.5)
abline(0, 1, lty = 2)

t_plot <- seq(-0.2, 1.2, length = 100)
plot(0, 0, xlim = c(-0.2, 1.2), ylim = range(y), type = 'n', xlab = 't', ylab = expression(theta(t)))
for (i in 1:n) lines(t[[i]] + res$w[i], y[[i]], col = rainbow(n)[i])
lines(t_plot, basis_fct(t_plot) %*% res$c, lwd = 2, lty = 2, col = 'red')
lines(t_plot, theta(t_plot), lwd = 2, lty = 2)

# Compare estimated variance parameters of warps

# True standard deviation
sigma * sqrt(scale)

# Standard deviation of true shifts 
sd(w)

# Estimated standard deviation
res$sigma * sqrt(res$warp_cov_par)

# Standard deviation of predicted warps
sd(res$w)
```


## Example 3: serial correlation and normally distributed shifts
We use the same data as in Example 2, but instad of iid normally distributed noise we now add a Matérn process to the observations. 
```{r}
# Generate shifts
scale <- 0.5
w <- rnorm(n, sd = sigma * sqrt(scale))

amp_cov <- make_cov_fct(Matern, noise = FALSE)


# Generate data with random shifts
y <- lapply(1:n, function(i) {theta(t[[i]] + w[i]) + sigma * t(chol(amp_cov(t[[i]], c(100, 0.3, 2)))) %*% rnorm(m)}) 

# Plot shifted curves
plot(0, 0, xlim = c(-0.2, 1.2), ylim = range(y), type = 'n', xlab = 't', ylab = expression(theta(t)))
for (i in 1:n) lines(t[[i]], y[[i]], col = rainbow(n)[i])
lines(t[[1]], theta(t[[1]]), lwd = 2, lty = 2)
```

Let us first ignore the uncertainty and use the same model as in Example 1.

```{r}
res_no_uncert <- pavpop(y, t, basis_fct, warp_fct, amp_cov = NULL, warp_cov = NULL, iter = c(1, 5))
```

And now compare to the correct specification of the model

```{r}
res <- pavpop(y, t, basis_fct, warp_fct, amp_cov = amp_cov, warp_cov = warp_cov, iter = c(5, 5))
```


We can not plot the results

```{r}
plot(w, res$w, xlab = 'true shifts', ylab = 'estimated shifts', pch = 19, col = 'green')
points(w, res_no_uncert$w, pch = 19, col = 'red', cex = 0.5)
abline(0, 1, lty = 2)

par(mfrow = c(1, 2))
plot(0, 0, xlim = c(-0.2, 1.2), ylim = range(y), type = 'n', xlab = 't', ylab = expression(theta(t)))
for (i in 1:n) lines(t[[i]] + res_no_uncert$w[i], y[[i]], col = rainbow(n)[i])
lines(t_plot, basis_fct(t_plot) %*% res_no_uncert$c, lwd = 2, lty = 2, col = 'blue')
lines(t_plot, theta(t_plot), lwd = 2, lty = 2)

plot(0, 0, xlim = c(-0.2, 1.2), ylim = range(y), type = 'n', xlab = 't', ylab = expression(theta(t)))
for (i in 1:n) lines(t[[i]] + res$w[i], y[[i]], col = rainbow(n)[i])
lines(t_plot, basis_fct(t_plot) %*% res$c, lwd = 2, lty = 2, col = 'blue')
lines(t_plot, theta(t_plot), lwd = 2, lty = 2)

# Compare estimated variance parameters of warps

# True standard deviation
sigma * sqrt(scale)

# Standard deviation of true shifts 
sd(w)

# Estimated standard deviation
res$sigma * sqrt(res$warp_cov_par)

# Standard deviation of predicted warps
sd(res$w)
```
